---
title: "p8105_hw3_pz2281"
author: "Peilin Zhou"
output: github_document
---

```{r}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%" #shrunk down the size 
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1
read in instacart data

```{r}
library(p8105.datasets)
data("instacart")
instacart
```

The instacart dataset contains `r nrow(instacart)` observations with `r ncol(instacart)` variables. Each row represents a product from an order. The data contains customer information with unique user_id as a variable and there are `r length(unique(instacart$user_id))` unique customers. There are total of `r length(unique(instacart$department))` departments and `r length(unique(instacart$aisles))` aisles. The products from each order_id are ordered by the add_to_cart_order from the customer.

```{r}
aisle_freq = instacart %>% 
  group_by(aisle) %>% 
  summarize(
    n_obs = n()
  )
aisle_freq %>% 
  slice_max(n_obs)
```





